The Hydrogen Epoch of Reionization Array Dish: Beam Pattern Measurements and Science Implications

### 2015-11-04 DishBeamMeasurements/paper/DishBeamMeasurementsPaper.tex

basics of 21cm cosmology

must use frequency structure to separate FGs from signal
a certain amount of freq structure is imprinted onto smooth sources due to the frequency-dependent sampling function of an interferometer
also by faraday rotation and the bandpass (citations)

leakage is due to bandpass, frequency flagging, finite bandwidth. can be mitigated by subtraction in the delay spectrum (CLEAN), giving up on wedge modes, or subtraction in image space with foreground modeling, likely will have to give up on wedge modes too unless have incredible calibration and imaging fidelity, as LOFAR is attempting. Large risk of signal loss and calibration-induced transfer of spectral structure from long baselines to short ones

hera is pursuing two relatively orthogonal power spectrum analyses (imaging and delay transform)

the other papers in this series quantify the degree of spectral structure induced by the dish bandpass in measurement and modeling, we focus in this section on foreground leakage due to the finite bandwidth as a function of the beam response at different delays in the sky. Nithya discuss this is detail nithya15,nithya15b

\citet{nithya15,nithya15b} 
for a power spectrum analysis, note that these tests test how leakage varies as a function of the brightness of emission at different delays, not leakage due to beam frequency dependence (ie, either pattern or overall gain)

\section{ORBCOMM Beam Mapping System Review}

@@

\section{Beam Modeling}

\subsection{Reference Antenna Modeling}
Our power ratio measurement relies upon an accurately modeled reference dipole. We are developing a more comprehensive model of the physical reference antenna and ground screen design, improving on that used in \citep{neben15}. As yet, this improved model yields incorrect results, so more debugging is needed.

\subsection{Dish and Feed Modeling}
\label{sec:dishmodels}

@@

\begin{enumerate}
	\item Airy pattern for a 14\,m aperture
	\item CST simulation (\texttt{hera-cst/HERA\_DISH\_paper\_feed\_cyl36\_140mhz\_Y\_healpix.fits})
	\item A separate simulation run by Dave with 68'' diameter feed backplane and 26'' skirt.
\end{enumerate}

\begin{figure*}[h]
\includegraphics[width=6.5in]{aerial.png}
\caption{Aerial photograph of the two reference dipoles deployed in Galford Meadow at the National Radio Astronomy Observatory--Green Bank 100\,m south of the HERA dish.}
\label{fig:aerial}
\end{figure*}

\section{Assessing Systematics with a Null Experiment}

@@

\begin{figure*}[h]
\includegraphics[width=6.5in]{null1_rel.png}
\caption{Top: ratio maps of the powers received by two reference dipoles separated by 50\,m, deployed 50\,m south of the HERA dish. Bottom: absolute beam power plots constructed by multiplying the relative power maps by a model beampattern. }
\label{fig:null1}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.5in]{null3_rel.png}
\caption{Same as the previous null experiment except the south reference dipole was moved 5\,m west. The systematics are largely unchanged.}
\label{fig:null3}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.5in]{null4_rel.png}
\caption{Same as the first null experiment except the reference dipoles are separated by 100\,m and deployed 100\,m south of the HERA dish. The systematics remain at the few percent level near zenith and increase to $\sim10\%$ farther out, with somewhat, but not entirely different angular structure. }
\label{fig:null4}
\end{figure*}

\section{Feed Measurements}

Accurate modeling of both the dish and feed will be essential for final dish modeling. To that end, we deploy the feed on the ground facing the sky (Figure \ref{fig:feedpic}), replacing the north-most reference antenna in the \texttt{null3} configuration, to isolate the feed power pattern.

Figure \ref{fig:feed0} shows the measured feed beams (top panel) and slices through the E and H planes (bottom panel). Due to intermittent connection problems to the DAQ computer, we collected only 157 high signal-to-background satellite passes over a few days. Still, the data agree very well with the CST model for $\theta<55^\circ$, outside of which the E plane response starts narrows compared to the model by up to 5\,dB at $\theta=70^\circ$. Still, the data are quite sparse in these low regions due to the connection problems, and so the usual quality control cuts are likely less effective than usual.

\begin{figure*}[h]
\includegraphics[width=5in]{feed.jpg}
\caption{Feed deployed on the ground, facing the sky, replacing the north-most reference antenna in the \texttt{null3} configuration. }
\label{fig:feedpic}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.5in]{feed0_abs_old_ref_model.png}
\caption{Measurements of the feed power pattern, deployed so it faces the sky. The measurements agree well with the model for $\theta<55^\circ$, but start to diverge at large zenith angles, possibly due to quality control problems due to very sparse beam sampling.}
\label{fig:feed0}
\end{figure*}

@@

\begin{figure*}[h]
\includegraphics[width=6.5in]{dish3_abs_old_ref_model.png}
\caption{Dish power pattern with the feed lowered 0.5\,m below nominal focus.}
\label{fig:dish3}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.5in]{dish1_abs_old_ref_model.png}
\caption{Dish power pattern with the feed at the nominal focus of 4.5\,m.}
\label{fig:dish1}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.5in]{dish2_abs_old_ref_model.png}
\caption{Dish power pattern with the feed raised 0.5\,m above nominal focus.}
\label{fig:dish2}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.5in]{dish4_abs_old_ref_model.png}
\caption{Dish power pattern with the feed raised to its maximum height of 0.8\,m above nominal focus.}
\label{fig:dish4}
\end{figure*}

@@

\begin{figure}[h]
\includegraphics[width=3.5in]{measbeam_raw.png}
\includegraphics[width=3.5in]{measbeam_interp.png}
\includegraphics[width=3.5in]{measbeam_interp_expcutoff.png}
\caption{We construct full sky beams as required for Section \ref{sec:foregroundleakage} by starting from the nominal focus beam (top left), interpolating over unobserved cells at $\theta\lesssim60^\circ$ and extrapolating to the horizon (top right), and applying a gaussian cutoff starting at $\theta=60^\circ$ with $\sigma=2.5^\circ$ (bottom left). These last two beams span the space of likely horizon responses. Circles denote $\theta=20^\circ,40^\circ,60^\circ,80^\circ$.}
\label{fig:interpbeams}
\end{figure}

\begin{figure}[h]
\includegraphics[width=3.5in]{ew_beams_slice.pdf}
\caption{We plot the interpolated/extrapolated measured nominal focus  beam itself (blue), and with the gaussian cutoff (cyan), along with various model beams to illustrate the varying horizon responses. }
\label{fig:interpbeamsslice}
\end{figure}

@@

\begin{figure*}[h]
\includegraphics[width=6.7in]{LST0deg_14m_42m_EWbaselines_dish1_blackman.png}
\includegraphics[width=6.7in]{LST60deg_14m_42m_EWbaselines_dish1_blackman.png}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspec}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{LST0deg_14m_42m_NEslantbaselines_dish1_blackman.png}
\includegraphics[width=6.7in]{LST0deg_14m_42m_SEslantbaselines_dish1_blackman.png}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ LST for NE (top pair) and SE (bottom pair) baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspec2}
\end{figure*}

%\section{Discussion}

%As expected, the Airy patter has the largest collecting area

%the beam might be narrower simply because more dish is illuminated, but not necessarily a good thing, because we want more taper at the edge

%comment on how this expt is still a work in progress

>>>>

%- basics of 21cm cosmology
    - shedding light on the unobserved Dark Ages and Epoch of
      Reionization when the first luminous sources formed and reionized
      the IGM
    - MWA, PAPER, LOFAR, GMRT power spectrum experiments
    - also global signal experiments like XX,YY,ZZ,
        - the big challenge is recovering the signal below 4-5 orders of
      magnitude brighter foregrounds and noise
    - naive requirement is collecting area
    - naive foreground removal is to subtract all low $k_\parallel$ modes

%- but instrumental frequency structure has been recognized as critical
  => wedge
    - of course the bandpass and intrinsic smooth freq structure of
      sources
    - but also the freq-dependent sampling of the IFO
    - sources at different zenith angles appear at different delays,
      and thus different $k_\parallel$, increasing linearly with baseline length
    - this gives rise to the “wedge” in cylindrically averaged power
      spectra ($k_\perp$, $k_\parallel$)

%- the beam affects the wedge
    - Nithya et al quantify the effects of antenna beam on the wedge
    - wider beams like the dipole have large response at large zenith
      angle, and thus large delays, wedge much more filled out, and
      much more emission close to the EOR window, at risk of leaking in
    - narrower beams like the airy have much narrower response, most
      received emission is from low delays (near $k_\parallel=0$)
    - but even there, emission is seen at the edge of the wedge,
      interpreted as an effective brightening of the horizon due to the
      very large solid angle it subtends 
    - this structure (lots of emission from zero delay, and peak at the
      horizon), was termed the pitchfork
    - highlights importance of beam characterization across the entire
      sky

%- the design of the HERA dish
    - small D, large N technique was enabled by moore’s law
    - even if processing hundreds to thousands of hours of data on
      order hundred of antennas is technically possible with parallel
      processing and advanced data quality metrics
    - first generation experiments are aiming for a first detection of
      the power spectrum, even more sensitivity will be needed to
      characterize its evolution with redshift and k, and test models
    - HERA has thus opted for larger elements, at the expense of a
      smaller FOV, but most of our fourier modes come from freq
      dimension anyway
    - a dish is preferred to a large phased array (a la SKA) as a
      simpler antenna, with fewer degrees of freedom and reduced
      potential for ant-to-ant variation (Neben et al)
    - of course, elevating a feed above a dish introduces potential for
      reflections, and thus a wigglier bandpass than that of the MWA or
      PAPER
    - The other papers in this series set a wiggliness spec and present
      measurements showing it meets the spec
    - here we focus on the angular response of the dish, ie, its beam
      pattern

%- how sensitive are we to accurate beam models?
    - what must we show about the HERA Dish to confirm that it will
      allow us to reach the science we want?
    - large collecting area is a given
    - minimal sidelobes and horizon response
    - how accurately must we verify the beam model? different analyses
      require different levels of beam modeling fidelity
    - HERA’s hexagonal grid permits redundant baseline calibration
      which is agnostic to the antenna beam, as long as all are the
      same. ant-to-ant variation adds noise to the redundant
      calibration, likely much smaller than that incurred in MWA sky
      model calibration 
    - the PAPER delay spectrum technique and 1D clean is even agnostic
      to ant-to-ant variation, as it cleans out all sky emission within
      the horizon limits on a per baseline basis
    - an imaging analysis (a la MWA) relies on foreground subtraction,
      and containment of residuals in the wedge, and the subtraction
      step is much more sensitive both to accurate modeling of the mean
      beam and to ant-to-ant variation (Neben et al)
    - the general questions we address in this paper
        - the questions are: does the main lobe shape agree with the
          model (need good model for imaging and FG subtraction)
        - are the sidelobe levels and horizon response sufficiently
          small to mitigate pitchfork effect and foreground leakage?

%- roadmap
    - we use orbcomm to measure the dish beam pattern and calculate
      collecting area, verify the focus
    - then we assess the science implications in delay space, comparing
      different models and power spectrum approaches

\subsection{HERA--Green Bank: A three-element prototype array}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{aerial.png}
%\caption{Aerial photograph of the two reference dipoles deployed in Galford Meadow at the National Radio Astronomy Observatory--Green Bank 100\,m south of the HERA dish.}
%\label{fig:aerial}
%\end{figure*}

\subsection{ORBCOMM Beam Mapping System Review}

@@

%Our power ratio measurement relies upon an accurately modeled reference dipole. We are developing a more comprehensive model of the physical reference antenna and ground screen design, improving on that used in \citep{neben15}. As yet, this improved model yields incorrect results, so more debugging is needed.

\section{Dish Design and Modeling}

\subsection{Design of the HERA Dish}

design principles
tradeoffs
description of the final design

\subsection{Dish Modeling}
\label{sec:dishmodels}

@@

\begin{enumerate}
	\item Airy pattern for a 14\,m aperture
	\item CST simulation (hera-cst/HERA\_DISH\_paper \_feed\_cyl36\_140mhz\_Y\_healpix.fits)
	\item A separate simulation run by Dave with 68'' diameter feed backplane and 26'' skirt.
\end{enumerate}

\section{Error Analysis}

@@

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{null1_rel.png}
%\caption{Top: ratio maps of the powers received by two reference dipoles separated by 50\,m, deployed 50\,m south of the HERA dish. Bottom: absolute beam power plots constructed by multiplying the relative power maps by a model beampattern. }
%\label{fig:null1}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{null3_rel.png}
%\caption{Same as the previous null experiment except the south reference dipole was moved 5\,m west. The systematics are largely unchanged.}
%\label{fig:null3}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{null4_rel.png}
%\caption{Same as the first null experiment except the reference dipoles are separated by 100\,m and deployed 100\,m south of the HERA dish. The systematics remain at the few percent level near zenith and increase to $\sim10\%$ farther out, with somewhat, but not entirely different angular structure. }
%\label{fig:null4}
%\end{figure*}

%\section{Feed Measurements}

%Accurate modeling of both the dish and feed will be essential for final dish modeling. To that end, we deploy the feed on the ground facing the sky (Figure \ref{fig:feedpic}), replacing the north-most reference antenna in the \texttt{null3} configuration, to isolate the feed power pattern.

%Figure \ref{fig:feed0} shows the measured feed beams (top panel) and slices through the E and H planes (bottom panel). Due to intermittent connection problems to the DAQ computer, we collected only 157 high signal-to-background satellite passes over a few days. Still, the data agree very well with the CST model for $\theta<55^\circ$, outside of which the E plane response starts narrows compared to the model by up to 5\,dB at $\theta=70^\circ$. Still, the data are quite sparse in these low regions due to the connection problems, and so the usual quality control cuts are likely less effective than usual.

%\begin{figure*}[h]
%\includegraphics[width=5in]{feed.jpg}
%\caption{Feed deployed on the ground, facing the sky, replacing the north-most reference antenna in the \texttt{null3} configuration. }
%\label{fig:feedpic}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{feed0_abs_old_ref_model.png}
%\caption{Measurements of the feed power pattern, deployed so it faces the sky. The measurements agree well with the model for $\theta<55^\circ$, but start to diverge at large zenith angles, possibly due to quality control problems due to very sparse beam sampling.}
%\label{fig:feed0}
%\end{figure*}

@@

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish3_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed lowered 0.5\,m below nominal focus.}
%\label{fig:dish3}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish1_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed at the nominal focus of 4.5\,m.}
%\label{fig:dish1}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish2_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed raised 0.5\,m above nominal focus.}
%\label{fig:dish2}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish4_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed raised to its maximum height of 0.8\,m above nominal focus.}
%\label{fig:dish4}
%\end{figure*}

@@

%\begin{figure}[h]
%\includegraphics[width=3.5in]{measbeam_raw.png}
%\includegraphics[width=3.5in]{measbeam_interp.png}
%\includegraphics[width=3.5in]{measbeam_interp_expcutoff.png}
%\caption{We construct full sky beams as required for Section \ref{sec:foregroundleakage} by starting from the nominal focus beam (top left), interpolating over unobserved cells at $\theta\lesssim60^\circ$ and extrapolating to the horizon (top right), and applying a gaussian cutoff starting at $\theta=60^\circ$ with $\sigma=2.5^\circ$ (bottom left). These last two beams span the space of likely horizon responses. Circles denote $\theta=20^\circ,40^\circ,60^\circ,80^\circ$.}
%\label{fig:interpbeams}
%\end{figure}

%\begin{figure}[h]
%\includegraphics[width=3.5in]{ew_beams_slice.pdf}
%\caption{We plot the interpolated/extrapolated measured nominal focus  beam itself (blue), and with the gaussian cutoff (cyan), along with various model beams to illustrate the varying horizon responses. }
%\label{fig:interpbeamsslice}
%\end{figure}

@@

%\begin{figure*}[h]
%\includegraphics[width=6.7in]{LST0deg_14m_42m_EWbaselines_dish1_blackman.png}
%\includegraphics[width=6.7in]{LST60deg_14m_42m_EWbaselines_dish1_blackman.png}
%\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
%\label{fig:delayspec}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.7in]{LST0deg_14m_42m_NEslantbaselines_dish1_blackman.png}
%\includegraphics[width=6.7in]{LST0deg_14m_42m_SEslantbaselines_dish1_blackman.png}
%\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ LST for NE (top pair) and SE (bottom pair) baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
%\label{fig:delayspec2}
%\end{figure*}

\section{Discussion}

add stuff here
